---
title: "Simulated (significant)"
output: html_notebook
---

```{r message = FALSE}
library(dplyr)
library(readr)
library(survey)
```

# Loading data

```{r message = FALSE}
data <- read_csv("~/Desktop/UPENN/Spring Term 2026/BDS5010_TA/LLM_simulateddata/Simulated.csv")
```

# Applying pre-registered exclusion criteria as in the preregistration.
## "Anyone who doesn't complete the study" (COULD NOT BE CHECKED)
## "Who takes longer than 10 minutes to complete the entire experiment" (COULD NOT BE CHECKED)
## "Who fails any one of the comprehension checks"

```{r}
# 349 participants were left (way more lenient with comprehension checks)
data <- data %>%
  filter(AUCTION_WINNER == "Participant 2")
```

```{r}
NAIU <- subset(data, CONDITION == "No AI x Utilitarian (new)")
NAIH <- subset(data, CONDITION == "No AI x Hedonic (new)")
AIU <- subset(data, CONDITION == "AI x Utilitarian (new)")
AIH <- subset(data, CONDITION == "AI x Hedonic (new)")
```

```{r}
NAIU2 <- NAIU %>%
  rename(
    age = AGE, 
    gender = GENDER, 
    acheck_1 = AI_MENTION_CHECK,
    PO_1 = PR_INCORP_SELF,
    PO_2 = PR_BELONGS_TO_ME,
    PO_3 = PR_CONNECTED,
    PO_4 = PR_CLOSENESS,
    PO_5 = PR_DIFFICULT_MINE,
    AI_1 = AI_IMPROVE_LIFE, 
    AI_2 = AI_IMPROVE_WORK, 
    AI_3 = AI_USE_FUTURE,
    AI_4 = AI_POSITIVE_HUMANITY,
    AI_5 = AI_THREAT_HUMANS,
    mcheck = PRODUCT_HEDONICITY_1_UTIL_7_HED,
    WTP = WTP_COINS_0_10, 
    MTurkID = MTURK_ID)

NAIH2 <- NAIH %>%
  rename(
    age = AGE, 
    gender = GENDER, 
    acheck_1 = AI_MENTION_CHECK,
    PO_1 = PR_INCORP_SELF,
    PO_2 = PR_BELONGS_TO_ME,
    PO_3 = PR_CONNECTED,
    PO_4 = PR_CLOSENESS,
    PO_5 = PR_DIFFICULT_MINE,
    AI_1 = AI_IMPROVE_LIFE, 
    AI_2 = AI_IMPROVE_WORK, 
    AI_3 = AI_USE_FUTURE,
    AI_4 = AI_POSITIVE_HUMANITY,
    AI_5 = AI_THREAT_HUMANS,
    mcheck = PRODUCT_HEDONICITY_1_UTIL_7_HED,
    WTP = WTP_COINS_0_10, 
    MTurkID = MTURK_ID)

AIU2 <- AIU %>%
  rename(
    age = AGE, 
    gender = GENDER, 
    acheck_1 = AI_MENTION_CHECK,
    PO_1 = PR_INCORP_SELF,
    PO_2 = PR_BELONGS_TO_ME,
    PO_3 = PR_CONNECTED,
    PO_4 = PR_CLOSENESS,
    PO_5 = PR_DIFFICULT_MINE,
    AI_1 = AI_IMPROVE_LIFE, 
    AI_2 = AI_IMPROVE_WORK, 
    AI_3 = AI_USE_FUTURE,
    AI_4 = AI_POSITIVE_HUMANITY,
    AI_5 = AI_THREAT_HUMANS,
    mcheck = PRODUCT_HEDONICITY_1_UTIL_7_HED,
    WTP = WTP_COINS_0_10, 
    MTurkID = MTURK_ID)

AIH2 <- AIH %>%
  rename(
    age = AGE, 
    gender = GENDER, 
    acheck_1 = AI_MENTION_CHECK,
    PO_1 = PR_INCORP_SELF,
    PO_2 = PR_BELONGS_TO_ME,
    PO_3 = PR_CONNECTED,
    PO_4 = PR_CLOSENESS,
    PO_5 = PR_DIFFICULT_MINE,
    AI_1 = AI_IMPROVE_LIFE, 
    AI_2 = AI_IMPROVE_WORK, 
    AI_3 = AI_USE_FUTURE,
    AI_4 = AI_POSITIVE_HUMANITY,
    AI_5 = AI_THREAT_HUMANS,
    mcheck = PRODUCT_HEDONICITY_1_UTIL_7_HED,
    WTP = WTP_COINS_0_10, 
    MTurkID = MTURK_ID)
```

```{r}
remove_straightliners <- function(df) {
  df %>%
    rowwise() %>%
    mutate(
      PO_unique = n_distinct(c_across(PO_1:PO_5)),
      AI_unique = n_distinct(c_across(AI_1:AI_5))
    ) %>%
    ungroup() %>%
    # Remove participants who straight-line PO or AI
    filter(PO_unique > 1 & AI_unique > 1) %>%
    dplyr::select(-PO_unique, -AI_unique)
}

NAIU2 <- remove_straightliners(NAIU2)
NAIH2 <- remove_straightliners(NAIH2)
AIU2  <- remove_straightliners(AIU2)
AIH2  <- remove_straightliners(AIH2)
```

# Preparing the dataset
## Selecting columns needed for the analysis

```{r}
NAIU3 <- NAIU2 %>%
  dplyr::select(PO_1, PO_2, PO_3, PO_4, PO_5, AI_1, AI_2, AI_3, AI_4, AI_5, WTP, gender, age, MTurkID)

NAIH3 <- NAIH2 %>%
  dplyr:: select(PO_1, PO_2, PO_3, PO_4, PO_5, AI_1, AI_2, AI_3, AI_4, AI_5, WTP, gender, age, MTurkID)

AIU3 <- AIU2 %>%
   dplyr::select(PO_1, PO_2, PO_3, PO_4, PO_5, AI_1, AI_2, AI_3, AI_4, AI_5, WTP, gender, age, MTurkID)

AIH3 <- AIH2 %>%
  dplyr:: select(PO_1, PO_2, PO_3, PO_4, PO_5, AI_1, AI_2, AI_3, AI_4, AI_5, WTP, gender, age, MTurkID)
```

## Adding conditions

```{r}
# AI = 0: NO AI
# AI = 1: AI
# HU = 0: UTILITARIAN
# HU = 1: HEDONIC

NAIU4 <- NAIU3 %>%
  mutate(
    AI = 0,
    HU = 0)

NAIH4 <- NAIH3 %>%
  mutate(
    AI = 0,
    HU = 1)

AIU4 <- AIU3 %>%
  mutate(
    AI = 1,
    HU = 0)

AIH4 <- AIH3 %>%
  mutate(
    AI = 1,
    HU = 1)
```

## Combining the dataset

```{r}
finaldata <- bind_rows(
  NAIU4,
  NAIH4,
  AIU4,
  AIH4
)

# Adding the interaction between AI & HU
## 0_0: NO AI & UTILITARIAN
## 1_0: AI & UTILITARIAN
## 0_1: NO AI & HEDONIC
## 1_1: AI & HEDONIC

# ChatGPT assisted with this code. 
finaldata <- finaldata %>%
  mutate(group = interaction(AI, HU, sep = "_"))
```

## Calculating variables needed for the analysis

```{r}
finaldata <- finaldata %>%
  mutate(PO_5 = 6 - PO_5) %>%
  mutate(
    PO_overall = rowMeans(
      cbind(PO_1, PO_2, PO_3, PO_4, AI_5)))

finaldata <- finaldata %>%
  mutate(AI_5 = 6 - AI_5) %>%
  mutate(
    AI_overall = rowMeans(
      cbind(AI_1, AI_2, AI_3, AI_4, AI_5)))
```

## Selecting columns needed for the analysis

```{r}
finaldata_2 <- finaldata %>%
  dplyr::select(PO_overall, AI_overall, WTP, gender, age, AI, HU, group, MTurkID)

# Checking for nature of variables included in the dataset
finaldata_2 <- finaldata_2 %>%
  mutate(
    WTP = as.numeric(WTP),
    PO_overall = as.numeric(PO_overall),
    AI_overall = as.numeric(AI_overall), 
    age = as.numeric(age), 
    gender = as.factor(gender)
  )
```

# Analyses
## Descriptive statistics

```{r}
# Descriptive statistics of focal variables
desc_stats <- finaldata_2 %>%
  summarise(
    n = n(),
    mean_WTP = mean(WTP),
    sd_WTP = sd(WTP),
    median_WTP = median(WTP),
    min_WTP = min(WTP),
    max_WTP = max(WTP),
    mean_PO = mean(PO_overall),
    sd_PO = sd(PO_overall),
    median_PO = median(PO_overall), 
    min_PO = min(PO_overall),
    max_PO = max(PO_overall),
    mean_AI = mean(AI_overall),
    median_AI = median(AI_overall),
    min_AI = min(AI_overall),
    max_AI = max(AI_overall),
    mean_age = mean(as.numeric(age)),
    median_age = median(as.numeric(age)),
    min_age = min(as.numeric(age)),
    max_age = max(as.numeric(age)))
desc_stats
```
## Visualization of WTP

```{r}
ggplot(finaldata_2, aes(x = WTP)) +
  geom_density(fill = "#B8A9E7", alpha = 0.5,
               color = "#3A2A6A", linewidth = 1.2) +
  labs(
    title = "Density Plot of Willingness to Pay (WTP)",
    x = "WTP",
    y = "Density"
  ) +
  theme_minimal()
```

## Regression

```{r}
model1 <- lm(WTP ~ AI + HU + PO_overall + age + gender + AI_overall, data = finaldata_2)
summary(model1)
```

## Barchart

```{r}
# Aggregate WTP by AI vs. No AI
bar_AI <- finaldata_2 %>%
  group_by(AI_label) %>%
  summarise(
    mean_WTP = mean(WTP, na.rm = TRUE),
    sd_WTP   = sd(WTP, na.rm = TRUE),
    n        = n(),
    se_WTP   = sd_WTP / sqrt(n)
  )

# Bar chart with error bars
ggplot(bar_AI, aes(x = AI_label, y = mean_WTP, fill = AI_label)) +
  geom_col(width = 0.6) +
  geom_errorbar(aes(ymin = mean_WTP - se_WTP,
                    ymax = mean_WTP + se_WTP),
                width = 0.15, linewidth = 0.8) +
  geom_text(aes(label = paste0("n = ", n),
                y = mean_WTP + se_WTP + 0.3),
            size = 4) +
  scale_fill_manual(values = c("No AI" = "#B8A9E7",
                               "AI"    = "#8172B3")) +
  coord_cartesian(ylim = c(0, max(bar_AI$mean_WTP + bar_AI$se_WTP) + 1)) +
  labs(
    title = "Mean WTP by AI Condition",
    x = "",
    y = "Mean WTP (0–10)"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")
```

## Kruskal-Wallis tests

```{r}
# No significant effect on WTP for AI vs.No AI + interaction AI condition and Product Type
kruskal.test(WTP ~ group, data = finaldata_2)
kruskal.test(WTP ~ AI, data = finaldata_2)
```

```{r}
# No significant results
kruskal.test(PO_overall ~ AI, data = finaldata_2)
kruskal.test(PO_overall ~ HU, data = finaldata_2)
kruskal.test(WTP ~ HU, data = finaldata_2)
```

### Visualization: Dumbbell Plot

```{r}
dumbbell_data <- finaldata_2 %>%
  group_by(HU, AI) %>%
  summarise(
    mean_WTP = mean(WTP, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    HU_label = ifelse(HU == 0, "Utilitarian", "Hedonic"),
    AI_label = ifelse(AI == 0, "No AI", "AI")
  )

ggplot(dumbbell_data,
       aes(x = AI_label,
           y = mean_WTP,
           group = HU_label,
           color = HU_label)) +
  
  geom_line(size = 1.2, alpha = 0.6) +
  geom_point(size = 4) +
  
  labs(
    title = "WTP by AI vs. No-AI Across Product Types",
    x = "AI Condition",
    y = "Mean WTP (0-10)",
    color = "Product Type"
  ) +
  
  scale_color_manual(values = c("purple", "black")) +
  
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 16)
  )
```

### Visualization: Boxplot AI vs. No AI

```{r}
finaldata_2$AI_label <- recode(finaldata_2$AI,
                               "0" = "No AI",
                               "1" = "AI")

n_counts <- finaldata_2 %>%
  group_by(AI_label) %>%
  summarise(n = n())

ggplot(finaldata_2, aes(x = AI_label, y = WTP, fill = AI_label)) +
  geom_boxplot() +
  scale_fill_manual(values = c("AI" = "#B8A9E7", "No AI" = "#B8A9E7")) +
  geom_text(data = n_counts,
            aes(label = paste0("n = ", n), y = max(finaldata_2$WTP) + 0.5),
            size = 5, color = "black") +
  labs(
    title = "WTP by AI vs. No AI",
    x = "",
    y = "WTP"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

### Visualization: Boxplot Four conditions

```{r}
# ChatGPT assisted with this code. 
finaldata_2$group_label <- recode(finaldata_2$group,
  "0_0" = "No AI\nUtilitarian",
  "1_0" = "AI\nUtilitarian",
  "0_1" = "No AI\nHedonic",
  "1_1" = "AI\nHedonic"
)

n_group <- finaldata_2 %>%
  group_by(group_label) %>%
  summarise(n = n())

ggplot(finaldata_2, aes(x = group_label, y = WTP, fill = group_label)) +
  geom_boxplot(alpha = 0.8, outlier.alpha = 0.3) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 4, color = "black") +
  geom_text(data = n_group,
            aes(label = paste0("n = ", n), y = max(finaldata_2$WTP) + 0.5),
            size = 4.5) +
  labs(
    title = "WTP Across Experimental Conditions",
    x = "Condition",
    y = "Willingness to Pay (WTP)"
  ) +
  scale_fill_manual(values = c("#B8A9E7", "#B8A9E7", "#B8A9E7", "#8172B3")) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 12))
```


